A Data-Driven Framework for Neural Field Modelling

### 2010-10-26 ltx/BrainIDEPoster/BrainIDE_Poster.tex
fixed the integral in the poster
simplified poster, bigger pictures

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{48in}
\setlength{\paperheight}{36in}
\setlength{\sepwid}{0.024\paperwidth}
\setlength{\onecolwid}{0.22\paperwidth}
\setlength{\twocolwid}{0.464\paperwidth}
\setlength{\threecolwid}{0.708\paperwidth}
\setlength{\topmargin}{-0.5in}
\usetheme{confposter}

%-----------------------------------------------------------
% Define colours (see beamerthemeconfposter.sty to change these colour definitions)
%-----------------------------------------------------------

@@

      \vskip2ex
% \begin{block}{Notation}
% 	\rmfamily{
% 	
% 	}
% \end{block}
      \begin{block}{Neural Field Model}
        \rmfamily{The model's temporal synaptic dynamics are described by 
		\begin{align}
			\label{SpikesToPotential} v\left( {\mathbf{r},t} \right) &= \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'} \\ 
			h(t) &= \eta(t)\exp{\left(-\zeta t\right)}, 
		\end{align}
		where $g(\mathbf{r},t)$ is the input firing rate, $\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function. Interactions between cortical populations are described by 
		\begin{equation}
			\label{RateBasedInteractions} g\left( \mathbf{r},t \right) = \int_\Omega {w\left( \left|\mathbf{r}-\mathbf{r}'\right| \right)f\left( v\left( \mathbf{r}',t \right) \right)\textrm{d}\mathbf{r}'}, 
		\end{equation}
		where $f(\cdot)$ is a sigmoidal firing rate function, $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing a cortical sheet or surface. Substituting equation~\ref{RateBasedInteractions} into equation~\ref{SpikesToPotential} gives the spatiotemporal model 
		\begin{equation}
			\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) =
			\int_{-\infty}^t 
			h\left(t - t'\right) \int_\Omega
			w\left(\mathbf{r},\mathbf{r}'\right) 
			f\left( v\left( \mathbf{r}',t' \right)\right)
			\textrm{d}\mathbf{r}'dt'.
		\end{equation}
The discrete time integro-difference equation form of the model is
\begin{alertblock}{Neural Field IDE}
\begin{equation}
	\label{DiscreteTimeModel} 
	v_{t+1}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r}'\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\textrm{d}\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
\end{alertblock}
where $T_s$ is the time step, $\xi = 1-T_s\zeta $ and $e_t(\mathbf{r})\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$.

@@

\begin{block}{State-Space Model}
	\begin{figure}	
	\begin{center}
	  \includegraphics[width=9in, scale = 10]{Figure2.eps}
	\end{center}
	\caption{{\bf Example of a one-dimensional field decomposition.} The dashed line depicts the field, the black solid lines shows the Gaussian basis functions and the vertical lines show the position and amplitude of the states.} 
	\label{fig:Figure2}	
	\end{figure}
	\rmfamily{To implement standard estimation techniques, we use a set of Gaussian basis functions to decompose the field \cite{Dewar2009}. The field decomposition is described by 
	\begin{equation}
		\label{DefFieldDecomp} v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t, 
	\end{equation}
	where $\mathbf{x}_t$ is the state vector that scales the field basis functions $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$. 
	The connectivity kernel can also be decomposed as 
	\begin{equation}\label{DefKernelDecomp}
		 w\left(\mathbf{r},\mathbf{r}'\right) =\boldsymbol{\psi}^\top\left(\mathbf{r},\mathbf{r}'\right) \boldsymbol{\theta},
	\end{equation}
	where $\boldsymbol{\psi}(\mathbf{r},\mathbf{r}')$ is a vector of Gaussian basis functions and $\boldsymbol{\theta}$ is a vector of scaling parameters. Now defining the matrices
	\begin{align}
		\boldsymbol{\Gamma} &\triangleq \int_\Omega {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi} ^{\top}\left(\mathbf{r}\right)\textrm{d}\mathbf{r}} \label{eq:DefGamma} \\ 
		\boldsymbol{\Psi}(\mathbf{r}') &\triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top}\label{eq:DefPsi} (\mathbf{r}'-\mathbf{r})\textrm{d}\mathbf{r}},
	\end{align}
	where $\Gamma$ is a $L\times L$ matrix and $\Psi$ is $L \times n_{\theta}$ matrix, where $L$ is the number of basis functions (states) and $n_{\theta}$ is the number of connectivity kernel basis functions. The state disturbance is defined as
	\begin{equation}\label{eq:Wt} 
		\mathbf{e}_t \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} )\textrm{d}\mathbf{r}},
	\end{equation}
	which is a zero mean normally distributed white noise process with covariance
	\begin{equation}
		\boldsymbol\Sigma_e =\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 
	\end{equation}
	By rearranging Eq.~\ref{DiscreteTimeModel} and substituting in Eqs. \ref{DefFieldDecomp}, \ref{eq:FieldBasisFunction}, \ref{eq:DefGamma}, \ref{eq:DefPsi} and \ref{eq:Wt} we get the following state-space model
	\begin{alertblock}{State-Space Model}
	\begin{align}
		\mathbf{x}_{t+1} &= \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \boldsymbol{\theta} + \xi\mathbf{x}_t + \mathbf{e}_t \\
		\mathbf{y}_t &= \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t.
		\end{align} 	
	\end{alertblock}
	where the observation matrix is 
	\begin{equation}
			\mathbf{C} = \left[
			\begin{array}{{ccc}} 
				c_{1,1} & \dots & c_{1,L} \\
				\vdots & \ddots & \vdots \\
				c_{N,1} & \dots & c_{N,L} 
			\end{array}
			\right] 
		\end{equation}
		and 
		\begin{equation}
			c_{i,j} = \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}')\textrm{d}\mathbf{r}'. 
		\end{equation}
	}
\end{block}

@@

      \begin{block}{Frequency Analysis}
		\begin{figure}
		\begin{center}
		\includegraphics[width=9in, scale = 10]{Figure4.eps}
		\end{center}
		\caption{{\bf Spatial frequency analysis of the neural field}. (\textbf{A}) The average (over time) power in dB of the spatial frequency of the neural field. (\textbf{B}) The average (over time) power in dB of the spatial frequency of the reconstructed neural field from the basis function decomposition.}
		\label{fig:Figure4}
		\end{figure}
		\rmfamily{Given a band-limited field, the distance between adjacent sensors, $\Delta_y$, must satisfy 
		\begin{equation}
			\label{eq:MinimumSensorDistance} \Delta_y \leq \frac{1}{2\rho_y\boldsymbol{\nu}_{c}}, 
		\end{equation}
		where $\rho_y \in \mathbb{R} \ge 1$ is an oversampling parameter and $\boldsymbol{\nu}_c$ is a cutoff frequency (typically taken as the -3~dB point). This condition must be satisfied to avoid spatial spectral aliasing effects when reconstructing the hidden dynamic field, $v_t(\mathbf{r})$, using the sampled observations, $\mathbf{y}_t$.

		In the same manner the minimum distance between basis functions must satisfy 
		\begin{equation}\label{eq:BasisFunctionSeparation}
			\Delta_{\phi} \leq \frac{1}{2\rho_{\phi}\boldsymbol{\nu}_{cy}}
		\end{equation}
		where $\rho_{\phi} \in \mathbb{R} \ge 1$ is an oversampling parameter to determine the basis function separation and $\boldsymbol{\nu}_{cy}$ is the cut-off frequency of the observed field.
		The field basis function widths can also be inferred using spectral considerations~\cite{Scerri2009}. The basis function width, $\sigma^2_{\phi}$, should be chosen to be
		\begin{equation}\label{eq:WidthCutOffRelationship}
		 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{cy}}^2},
		\end{equation}
		where
		\begin{equation}\label{eq:WidthFrequencyRelationship}
		 \sigma^2_{\nu_{cy}}= \frac{2\boldsymbol\nu_{cy}^\top \boldsymbol\nu_{cy}}{\ln2}.
		\end{equation}}
	\end{block}
	% \vskip2ex
	\begin{block}{Estimation}
		\rmfamily{The estimation procedure consists of a two-stage iterative algorithm incorporating the unscented Rauch-Tung-Striebel smoother for state estimation and a least squares algorithm for parameter estimation.} 
	\end{block}	
	\begin{block}{References}
        \small{\rmfamily{\begin{thebibliography}{99}
\bibitem{Deco2008} G. Deco, V. Jirsa, P. Robinson, M. Breakspear, K. Friston. PLoS Computational Biology. \textbf{4} (2008).
\bibitem{Dewar2009} M. Dewar, K. Scerri, V. Kadirkamanathan. IEEE Transactions on Signal Processing. \textbf{57} (2009) 83-91.
\bibitem{Scerri2009} K. Scerri, M. Dewar, V. Kadirkamanathan. IEEE Trans. on Signal Processing. \textbf{57} (2009) 482-492.
        \end{thebibliography}}}	\end{block}
\end{column}

>>>>

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{48in}
\setlength{\paperheight}{36in}
\setlength{\sepwid}{0.024\paperwidth}
\setlength{\onecolwid}{0.22\paperwidth}
\setlength{\twocolwid}{0.464\paperwidth}
\setlength{\threecolwid}{0.708\paperwidth}
\setlength{\topmargin}{-0.5in}
\usetheme{confposter}
\usepackage{exscale}
%-----------------------------------------------------------
% Define colours (see beamerthemeconfposter.sty to change these colour definitions)
%-----------------------------------------------------------

@@

      \vskip2ex
% \begin{block}{Notation}
% 	\rmfamily{
% 	
% 	}
% \end{block}
      \begin{block}{Neural Field Model}
        \rmfamily{The model's temporal synaptic dynamics are described by 
		\begin{align}
			\label{SpikesToPotential} v\left( {\mathbf{r},t} \right) &= \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'} \\ 
			h(t) &= \eta(t)\exp{\left(-\zeta t\right)}, 
		\end{align}
		where $g(\mathbf{r},t)$ is the input firing rate, $\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function. Interactions between cortical populations are described by 
		\begin{equation}
			\label{RateBasedInteractions} g\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)\textrm{d}\mathbf{r}'}, 
		\end{equation}
		where $f(\cdot)$ is a sigmoidal firing rate function, $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing a cortical sheet or surface. Substituting equation~\ref{RateBasedInteractions} into equation~\ref{SpikesToPotential} gives the spatiotemporal model 
		\begin{equation}
			\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) =
			\int_{-\infty}^t 
			h\left(t - t'\right) \int_\Omega
			w\left(\mathbf{r}-\mathbf{r}'\right) 
			f\left( v\left( \mathbf{r}',t' \right)\right)
			\textrm{d}\mathbf{r}'dt'.
		\end{equation}
The discrete time integro-difference equation form of the model is
\begin{alertblock}{Neural Field IDE}
\begin{equation}
	\label{DiscreteTimeModel} 
	v_{t+1}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r}-\mathbf{r}'\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\textrm{d}\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
\end{alertblock}
where $T_s$ is the time step, $\xi = 1-T_s\zeta $ and $e_t(\mathbf{r})\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$.

@@

\begin{block}{State-Space Model}
	\begin{figure}	
	\begin{center}
	  \includegraphics[width=9in, scale = 10]{Figure2.eps}
	\end{center}
	\caption{{\bf Example of a one-dimensional field decomposition.} The dashed line depicts the field, the black solid lines shows the Gaussian basis functions and the vertical lines show the position and amplitude of the states.} 
	\label{fig:Figure2}	
	\end{figure}
	\rmfamily{To implement standard estimation techniques, we use a set of Gaussian basis functions to decompose the field \cite{Dewar2009}. The field decomposition is described by 
	\begin{equation}
		\label{DefFieldDecomp} v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t, 
	\end{equation}
	where $\mathbf{x}_t$ is the state vector that scales the field basis functions $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$. 
	The connectivity kernel can also be decomposed as 
	\begin{equation}\label{DefKernelDecomp}
		 w\left(\mathbf{r}-\mathbf{r}'\right) =\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right) \boldsymbol{\theta},
	\end{equation}
	where $\boldsymbol{\psi}(\mathbf{r}-\mathbf{r}')$ is a vector of Gaussian basis functions and $\boldsymbol{\theta}$ is a vector of scaling parameters. Now defining the matrices
	\begin{align}
		\boldsymbol{\Gamma} &\triangleq \int_\Omega {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi} ^{\top}\left(\mathbf{r}\right)\textrm{d}\mathbf{r}} \label{eq:DefGamma} \\ 
		\boldsymbol{\Psi}(\mathbf{r}') &\triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top}\label{eq:DefPsi} (\mathbf{r}'-\mathbf{r})\textrm{d}\mathbf{r}},
	\end{align}
	where $\Gamma$ is a $L\times L$ matrix and $\Psi$ is $L \times n_{\theta}$ matrix, where $L$ is the number of basis functions (states) and $n_{\theta}$ is the number of connectivity kernel basis functions. The state disturbance is defined as
	\begin{equation}\label{eq:Wt} 
		\mathbf{e}_t \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} )\textrm{d}\mathbf{r}},
	\end{equation}
	which is a zero mean normally distributed white noise process with covariance
	\begin{equation}
		\boldsymbol\Sigma_e =\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 
	\end{equation}
	By rearranging Eq.~\ref{DiscreteTimeModel} and substituting in Eqs. \ref{DefFieldDecomp}, \ref{eq:FieldBasisFunction}, \ref{eq:DefGamma}, \ref{eq:DefPsi} and \ref{eq:Wt} we get the following state-space model
	\begin{alertblock}{State-Space Model}
	\begin{align}
		\mathbf{x}_{t+1} &= \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \boldsymbol{\theta} + \xi\mathbf{x}_t + \mathbf{e}_t \\
		\mathbf{y}_t &= \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t.
		\end{align} 	
	\end{alertblock}
	where the observation matrix is 
	\begin{align}
		\mathbf{C} &= \left[
		\begin{array}{{ccc}} 
			c_{1,1} & \dots & c_{1,L} \\
			\vdots & \ddots & \vdots \\
			c_{N,1} & \dots & c_{N,L} 
		\end{array}
		\right], \ \	c_{i,j} = \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}')\textrm{d}\mathbf{r}'. 
		\end{align}
	}
\end{block}

@@

      \begin{block}{Frequency Analysis}
		\begin{figure}
		\begin{center}
		\includegraphics[width=9in, scale = 10]{Figure4.eps}
		\end{center}
		\caption{{\bf Spatial frequency analysis of the neural field}. (\textbf{A}) The average (over time) power in dB of the spatial frequency of the neural field. (\textbf{B}) The average (over time) power in dB of the spatial frequency of the reconstructed neural field from the basis function decomposition.}
		\label{fig:Figure4}
		\end{figure}
		\rmfamily{Given a band-limited field, the distance between electrodes, $\Delta_y$, must satisfy 
		\begin{equation}
			\label{eq:MinimumSensorDistance} \Delta_y \leq \frac{1}{2\rho_y\boldsymbol{\nu}_{c}}, 
		\end{equation}
		where $\rho_y \in \mathbb{R} \ge 1$ is an oversampling parameter and $\boldsymbol{\nu}_c$ is a cutoff frequency (typically taken as the -3~dB point). This condition must be satisfied to avoid spatial spectral aliasing effects when reconstructing the hidden dynamic field, $v_t(\mathbf{r})$, using the sampled observations, $\mathbf{y}_t$.

		In the same manner the minimum distance between basis functions must satisfy 
		\begin{equation}\label{eq:BasisFunctionSeparation}
			\Delta_{\phi} \leq \frac{1}{2\rho_{\phi}\boldsymbol{\nu}_{cy}}
		\end{equation}
		where $\rho_{\phi} \in \mathbb{R} \ge 1$ is an oversampling parameter to determine the basis function separation and $\boldsymbol{\nu}_{cy}$ is the cut-off frequency of the observed field.
		The field basis function widths can also be inferred using spectral considerations~\cite{Scerri2009}. The basis function width, $\sigma^2_{\phi}$, should be chosen to be
		\begin{align}\label{eq:WidthCutOffRelationship}
		 \sigma^2_{\phi} &= \frac{1}{\pi^2\sigma_{\nu_{cy}}^2} \\
		 \sigma^2_{\nu_{cy}} &= \frac{2\boldsymbol\nu_{cy}^\top \boldsymbol\nu_{cy}}{\ln2}.
		\end{align}}
	\end{block}
	% \vskip2ex
	\begin{block}{Estimation}
		\rmfamily{The estimation procedure consists of a two-stage iterative algorithm incorporating the unscented Rauch-Tung-Striebel smoother for state estimation and a least squares algorithm for parameter estimation.} 
	\end{block}	
	\begin{block}{References}
        \small{\rmfamily{\begin{thebibliography}{99}
\bibitem{Deco2008} G. Deco, V. Jirsa, P. Robinson, M. Breakspear, K. Friston. PLoS Computational Biology. \textbf{4} (2008).
\bibitem{Dewar2009} M. Dewar, K. Scerri, V. Kadirkamanathan. IEEE Transactions on Signal Processing. \textbf{57} (2009) 83-91.
\bibitem{Scerri2009} K. Scerri, M. Dewar, V. Kadirkamanathan. IEEE Trans. on Signal Processing. \textbf{57} (2009) 482-492.
        \end{thebibliography}}}	\end{block}
\end{column}
